Clouds and Radiation in a mock-Walker Circulation

### 2019-10-29 Silvers_WalkerCell.tex
relatively minor text edits to Silvers_WalkerCell.tex.  Addition of two .bib files that contain references to use in Silvers_WalkerCell.tex

  a forced compositing of the data (\cite{Bretherton_etal_2005} Bretherton et al. 2005, Muller and Held, 2012).}

@@

We perform a series of sensitivity experiments that highlight the different ways in which these experiments can equilibrate.  To 
highlight both the robust features of the simulations and the sensitive nature of the distribution of the precipitation maximum 
we test the impact of convective parameterization, 
longwave radiation interactions with clouds, domain size, 
and the resolution, or grid-spacing.    Muller and Held, 2012 found that self-aggregation was sensitive to both domain size and resolution
because of the sensitivity of the low-level clouds to the longwave radiative cooling.

@@

To develop the model configuration used for these experiments we started with the same code base as that of the 
recently developed atmospheric global climate model AM4.0 
(Zhao et al. a,b).  This includes the GFDL finite-volume cubed-sphere dynamical core FV3 (Harris and Lin, 2013) 
which can solve either the hydrostatic primitive equations or the nonhydrostatic fully compressible Euler equations.  
The default AM4.0 physics we use includes interactive radiation, parameterized deep- and shallow-convection, 
a large-scale cloud scheme, and a boundary layer 
scheme as described in Zhao et al. a,b and the references therein.  The prognostic moisture variables are the specific 
humidity (q), liquid (ql) and frozen water (qi), and cloud fraction.  The top of the model domain is at 1 hPa, with 33 vertical 
levels and a sponge layer extending downward to 8 hPa.  The kilometer of atmosphere just above the surface is resolved by 
8 model levels.  The changes made to the default AM4.0 physics configuration are the following.  The cloud-aerosol 
interactions were turned off to focus on the interaction between clouds, radiation, and the circulation.  The gravity wave drag 
parameterization was turned off 
%(do_cg_drag set to .f.; Alexander-Dunkerton gravity wave drag) 
in order to reduce large oscillations which developed in the horizontal wind field near the top of the model domain.  
The convection, radiation, large-scale cloud, microphysics, and turbulence parameterizations all remain the same 
as in AM4.0.   Thus for the experiments with convective parameterization (grid-spacing of 25km and 100km), the 
physics are very similar to those of AM4.0.  This configuration can also be compared to the simulations described in 
Popp and Silvers, 2017, which used a developmental version of AM4 with almost the same changes to the physics
schemes.

@@

The mock-Walker Circulation that emerges from these simulations is shown in Figures \ref{fig:rh_psi_P25vsE25} and 
\ref{fig:precip_vertvel} to be characterized by a strong overturning circulation with precipitation focused over the warmer 
SSTs and a humid boundary layer across the full length of the domain.   For the GCM-like configuration (P25, left panels 
of Figure \ref{fig:rh_psi_P25vsE25}) with fully parameterized convection and with the longwave CRE turned on, there is 
a strong overturning circulation that results in a humid, condensate loaded troposphere near the middle of the domain 
and a dry ($< 10\% RH$) subsidence region with almost no condensate above $800 \rm{hPa}$.    
To illustrate some of the sensitivities to convection and the interaction between clouds, radiation, and the large-scale 
circulation we compare the P25 experiment with analogous experiments in which the longwave CRE is turned 
off (P25 LWCRE off, middle 
panels of Figure \ref{fig:rh_psi_P25vsE25}) and in which the convection is made explicit by turning off the convective 
parameterization (E25, right panels of Figure \ref{fig:rh_psi_P25vsE25}).  
The mass streamfunction and vertical velocity both show the P25 experiment has a stronger overturning circulation 
than either E25 or P25 LWCRE off.  P25 also has higher humidity in the deep convective region, and lower humidity in 
the subsidence region over the 
cooler SST (Figure \ref{fig:rh_psi_P25vsE25}) than either E25 or P25 LWCRE off.  Averaged over the full domain, 
the parameterized convection leads to a dryer atmosphere with less 
condensate (both liquid and ice) and with weaker radiative cooling and condensational heating (below 800 hPa).  
However, in the updraft region the convective parameterization results in more moisture aloft.

@@

One of the most prominent features of our GCM-like (particularly P25) simulations is an asymmetry relative to the 
symmetric SST distribution in both the time-dependent and 
steady-state solutions (Figures \ref{fig:rh_psi_P25vsE25},  \ref{fig:precip_vertvel}, \ref{fig:domdep} and  \ref{fig:conv_vs_ls}).  
This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$. 
The steady-state precipitation maximum is located  not over the warmest SST but is
shifted to slightly cooler temperatures.  The asymmetry is apparent in the vertical velocity, mass circulation, relative humidity, 
specific humidity, and radiative heating.  
In the Hovmoller diagrams (Figures \ref{fig:domdep} and \ref{fig:domdep_lwoff}), the precipitation appears to be averse to 
residing over the SST maximum.  
For the P25 case shown in Figure \ref{fig:rh_psi_P25vsE25} a strong ($1 \, \rm{m \, s^-1}$) domain
mean shear develops above about 500 \textit{m} which shifts the precipitation and circulation off center for years at a time.
When the convective parameterization is turned off (E25), the overturning circulation becomes weaker and 
broader (w, and psi), and the precipitation, cloud fields, and 
circulation all consistently reside over the SST maximum (for about 1 year before going haywire).   
While the parameterized convection plays a large role in driving this asymmetry it appears to be not entirely a 
result of the convective parameterization, but also due to an interaction between the convective 
parameterization and the LW CRE.   The degree to which this asymmetry influences the comparison with other experiments is 
unclear.    Note: check the wind shear for P25large, P100small, and P100large (E25 has less shear than E1 or E2).

Complex patterns of precipitation over a fixed sinusoidal or Gaussian SST distribution have been noted many times in 
previous literature (Bretherton et al. 2006, Kuang and Wofsy, Ming Zhao (personal communication, see Jeevanjee et al. 2018))
but the irregularities have 
tended to be symmetric about the SST maximum.  This is broadly consistent with our simulations when the convection is 
entirely explicit (E25, E2, and E1, discussed further in section 4).

%%%%%
% section 3.1 results
\begin{figure}
  \centering
      \includegraphics[width=0.96\columnwidth]{walkerfigs/rh_psi_cond_25km_6pan.eps}
      \caption{Comparison of the steady state relative humidity (shading of top panels) and mass 
      streamfunction (contours) for the experiments
      with a grid-spacing of 25km.  The lower panels show the total (liquid + ice) condensate (g/kg).  
      An experiment with fully parameterized convection (P25) is on the left, the same 
      experiment but with the LWCRE off is in the center panel, and the experiment with LWCRE on, but the convective
      parameterization turned off is on the right (E25).  All three panels use the same contour interval for the mass 
      streamfunction (kg/s).}
  \label{fig:rh_psi_P25vsE25}
\end{figure}

@@

Previous studies have shown that cloud radiative effects act to strengthen and contract an overturning circulation 
(e.g. Popp and Silvers, 2017, HH, and others (Muller and Held?)).  This was shown to be due to an increased low-level flux
of moist static energy into the convective regions.  While our mock-Walker circulation is distinct from radiative
convective equilibrium and the resulting convective self-aggregation, there are obvious similarities between our 
region of persistent deep convection and a state of aggregation.   One of the simplest measures of convective 
aggregation and the large-scale circulation is the fraction of the domain in which the air is subsiding, the 
subsidence fraction (SF).  As convection becomes more organized, or aggregated, SF will increase.  We expect 
for an overturning circulation that a contraction of the convective region would result in a larger subsidence 
fraction.  This is precisely what we see in Table 2.  For each of our experiments, the cases with LWCRE on
have a larger SF (with the exception of E1, for which SF is constant).  Our prescribed SST warm patch ensures that our simulations will 
be `aggregated' to some degree.  However, the subsidence fraction is a useful metric even in this case which 
includes an additional constraint than RCE.  Given identical SSTs, the variations in SF give a measure of variability
driven entirely by the atmosphere.

@@

Because of the tight constraints that connect the domain mean precipitation, atmospheric condensational heating, and 
the total radiative cooling, the time evolution of the precipitation serves as a useful measure of whether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that this equilibrium is reached after about 30 
days for the E2, and E1 simulations, and after about 50 days for the P25 and P100 simulations.   After the initial adjustments
the simulations all oscillate about mean precipitation values which tend to increase with the grid-spacing 
(Table 2).   The fact that E1 and E2 reach equilibrium sooner than E25, P25, or P100 and that the period of oscillation 
about the domain mean precipitation is smaller helps to justify the 6 month simulation times for E1 and E2.  
%The period of oscillation is much larger for the GCM like simulations than it is for the CRM like oscillations.
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   
%The experiments with a higher resolution have a lower value of domain
%mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  
This can be understood as a consequence of the differences in upper level cloud fraction and the surface energy budget 
and will be discussed further in a later section.

@@

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are 
two of the most fundamental choices that must be made.   One the goals of this study was to explore the results of 
changing grid-spacing over a wide range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies 
(e.g. Bretherton et al., 2006; Muller and Held, 2012; Jeevanjee and Romps, 2013; sens of what? Patrizio and Randall? Silvers?) have 
documented sensitivities to domain size.   The analysis of the previous, and of the next sections focus on results from 
experiments using a domain with a long domain length of $4000 \, \rm{km}$.  However we do find interesting 
sensitivities to the domain size that are noted here.

The evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, 
and the effect of the longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a domain with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with the idea that CREs strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation changes more as a function of domain size than it does as a function of grid-spacing. 
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitation.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.  As previously notes, these
GCM-like simulations equilibrate after approximately 50 days.  However, it is interesting to note a dramatic change in 
the distribution of precipitation on the $4000\, \rm{km}$ domain simulations after almost 2 years.   The domain mean 
precipitation does not significantly change in these cases, only the spatial structure.

@@

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the small domains, relative to a control domain size that was 4x
larger.   In contrast, here larger domains have larger precipitation rates (Table 2).  For the LWCRE off cases, we also find a more 
focused ascent region in the smaller domain.   But when the LWCRE is on, the extent of the ascent region changes little between P25 
and P25L but for P100 and P100L the ascent region is much more focuses on the large domain.    
How do our low-level clouds over the colder SSTs compare with the findings of B06 (check the condensate for small vs. large 
domains)?

% Fraction of Precip due to large-scale precip for large/small experiments with a 25km resolution
\begin{figure}
  \centering
      \includegraphics[width=0.8\columnwidth]{walkerfigs/Conv_vs_LS_hist.eps}
  \caption{Precipitation that is due to the large-scale(blue) cloud scheme and to the convective parameterization (black).}
  \label{fig:conv_vs_ls}
\end{figure}

The domain mean total precipitation is constrained by the radiative cooling of the atmosphere.  However, in 
models with the convection parameterized the 
total precipitation is composed of precipitation from the convection scheme and the large-scale cloud scheme. 
While the relative contribution is not well constrained, it is thought to influence the atmospheric state (Held et al., what part of the state?). 
The distribution of convective and large-scale precipitation can often indicate how the condensational heating
in a GCM is being distributed among the parameterizations, and what is triggering the precipitation.  
Precipitation from each of these two components is
shown in Figure \ref{fig:conv_vs_ls} as a function of both resolution and domain size.  
In the regions of large-scale ascent, most of the 
precipitation derives from the large-scale cloud scheme.  We also see that the LW CRE (thick lines) dramatically increases
the large-scale precipitation.  The LW CRE has a much smaller effect on the magnitude of the convective precipitation 
but does act to spatially concentrate it.  With the single exception 
of the P100L LW CRE off experiment, the convective precipitation produces relatively little of the total precipitation for 
both resolutions and on both domains.  The dramatic dependence on domain size of the precipitation field that is 
seen in Figure \ref{fig:domdep_lwoff} corresponds to decrease in the large-scale precipitation of about 65\% in P25L case
and an almost complete elimination of the large-scale precipitation in the P100L case.

@@

% this paragraph should perhaps be moved to the introduction...
It is clear that the overturning circulation in the tropical Pacific encompasses dynamical motions at a host of scales and that the 
different cloud types interact with the circulation on scales ranging from meters to thousands of km's.  
The tropical Pacific also serves as an ideal location to study interactions between clouds and the 
circulation because it includes strong overturning circulations, abundant shallow cumulus, congestus, and cumulonimbus 
clouds (Johnson et al. 1999), as well as stratocumulus cloud decks along the eastern extremities of the basin.
Can we hope to simulate all of this in a single idealized modeling context?    
Our initial motivation for using the GFDL AM4 model on a doubly-periodic domain was to simulate this region 
%that is rich in phenomena 
in the context of both a GCM and a CRM in the hopes that resolving more of the turbulent motions and circulations
would help us to better understand the physics and the mechanisms which are at work in the cloud-circulation
interactions of the tropical Pacific.    
The dynamical core of the current generation GFDL models is ideally suited to this task because it can easily 
be used over a wide range of resolutions and can solve either the hydrostatic or nonhydrostatic 
governing equations.

\begin{figure}
  \centering
      \includegraphics[width=0.96\columnwidth]{walkerfigs/rh_psi_cond_E25E2E1.eps}
      \caption{The steady state relative humidity (shading) and mass streamfunction (black contours).  
      Shown is E25 (left), E2(center), and E1(right).  All panels use the same contour interval for the 
      mass streamfunction (kg/s).}
  \label{fig:rh_psi_P25E2E1}
\end{figure}

@@

In this section we focus on simulations with grid-spacing of 1km, 2km, and 25km.   
A glance at Figure \ref{fig:hov_4mods_6mn} shows notable differences in the precipitation structure that result from the 
overturning circulation at different resolutions.  Shown are 180 days of precipitation from the P25, E25, E2, and 
E1 simulations.  As the 
resolution increases towards $1 \, \rm{km}$ the distribution of precipitation becomes more regular and consistently 
centered over the SST maxima in the middle of the domain.  Both the P25 and E25 simulations show more 
variability at later times, compared to these first 180 days.  The simulations with explicit convection at resolutions 
typical of cloud-resolving models (E2, E1) show little aversion to the precipitation maximum occurring over 
the maximum in SST and relative to the P100, P25 and E25 simulations 
they are able to maintain a smoother distribution of  precipitation values over a broader range of SST values.

@@

>>>>

  a forced compositing of the data  
  \cite{Bretherton2005, Muller2012}.
  %Bretherton et al. 2005, Muller and Held, 2012)
  }

@@

We perform a series of sensitivity experiments that highlight the different ways in which these experiments can equilibrate.  To 
highlight both the robust features of the simulations and the sensitive nature of the distribution of the precipitation maximum 
we test the impact of convective parameterization, 
longwave radiation interactions with clouds, domain size, 
and the resolution, or grid-spacing.    Muller and Held, 2012 found that self-aggregation was sensitive to both domain size and resolution
because of the sensitivity of the low-level clouds to these parameters.  % the longwave radiative cooling.

@@

To develop the model configuration used for these experiments we started with the same code base as that of the 
recently developed atmospheric global climate model AM4.0 \cite{Zhao_etal18a, Zhao_etal18b}.
%(Zhao et al. a,b).  
This includes the GFDL finite-volume cubed-sphere dynamical core FV3 (Harris and Lin, 2013) 
which can solve either the hydrostatic primitive equations or the nonhydrostatic fully compressible Euler equations.  
The default AM4.0 physics we use includes interactive radiation, parameterized deep- and shallow-convection, 
a large-scale cloud scheme, and a boundary layer 
scheme as described in Zhao et al. a,b and the references therein.  The prognostic moisture variables are the specific 
humidity (q), liquid (ql) and frozen water (qi), and cloud fraction.  The top of the model domain is at 1 hPa, with 33 vertical 
levels and a sponge layer extending downward to 8 hPa.  The kilometer of atmosphere just above the surface is resolved by 
8 model levels.  The changes made to the default AM4.0 physics configuration are the following.  The cloud-aerosol 
interactions were turned off to focus on the interaction between clouds, radiation, and the circulation.  The gravity wave drag 
parameterization was turned off 
%(do_cg_drag set to .f.; Alexander-Dunkerton gravity wave drag) 
in order to reduce large oscillations which developed in the horizontal wind field near the top of the model domain.  
The convection, radiation, large-scale cloud, microphysics, and turbulence parameterizations all remain the same 
as in AM4.0.   Thus for the experiments with convective parameterization (grid-spacing of 25km and 100km), the 
physics are very similar to those of AM4.0.  This configuration can also be compared to the simulations described in 
\cite{Popp2017}, %Popp and Silvers, 2017, 
which used a developmental version of AM4 with almost the same changes to the physics
schemes.

@@

The mock-Walker Circulation that emerges from these simulations is shown in Figures \ref{fig:rh_psi_P25vsE25} and 
\ref{fig:precip_vertvel} to be characterized by a strong overturning circulation with precipitation focused over the warmer 
SSTs and a humid boundary layer across the full length of the domain.   For the GCM-like configuration (P25, left panels 
of Figure \ref{fig:rh_psi_P25vsE25}) with fully parameterized convection and with the longwave CRE turned on, there is 
a strong overturning circulation that results in a humid, condensate loaded troposphere near the middle of the domain 
and a dry ($< 10\% RH$) subsidence region with almost no condensate above $800 \rm{hPa}$.    
This atmospheric state has the same fundamental characteristics as the Walker circulation, the Hadley Cell, and 
experiments of radiative convective equilibrium which equilibrate to a state with deep-overturning circulations and 
convective aggregation.

To illustrate some of the sensitivities to convection and the interaction between clouds, radiation, and the large-scale 
circulation we compare the P25 experiment with analogous experiments in which the longwave CRE is turned 
off (P25 LWCRE off, middle 
panels of Figure \ref{fig:rh_psi_P25vsE25}) and in which the convection is made explicit by turning off the convective 
parameterization (E25, right panels of Figure \ref{fig:rh_psi_P25vsE25}).  
The mass streamfunction and vertical velocity both show the P25 experiment has a stronger overturning circulation 
than either E25 or P25 LWCRE off.  P25 also has higher humidity in the deep convective region, and lower humidity in 
the subsidence region over the 
cooler SST (Figure \ref{fig:rh_psi_P25vsE25}) than either E25 or P25 LWCRE off.  Averaged over the full domain, 
the parameterized convection leads to a dryer atmosphere with less 
condensate (both liquid and ice) and with weaker radiative cooling and condensational heating (below 800 hPa).  
However, in the updraft region the convective parameterization results in more moisture aloft.

@@

One of the most prominent features of our GCM-like (particularly P25) simulations is an asymmetry relative to the 
symmetric SST distribution in both the time-dependent and 
steady-state solutions (Figures \ref{fig:rh_psi_P25vsE25},  \ref{fig:precip_vertvel}, \ref{fig:domdep} and  \ref{fig:conv_vs_ls}).  
This asymmetry is also present in P25 experiments on larger domains and at a resolution of $100 \, \rm{km}$. 
The steady-state precipitation maximum is located  not over the warmest SST but is
shifted to slightly cooler temperatures.  The asymmetry is apparent in the vertical velocity, mass circulation, relative humidity, 
specific humidity, and radiative heating.  
In the Hovmoller diagrams (Figures \ref{fig:domdep} and \ref{fig:domdep_lwoff}), the precipitation appears to be averse to 
residing over the SST maximum.  
For the P25 case shown in Figure \ref{fig:rh_psi_P25vsE25} a strong ($1 \, \rm{m \, s^-1}$) domain
mean shear develops above about 500 \textit{m} which shifts the precipitation and circulation off center for years at a time.
When the convective parameterization is turned off (E25), the overturning circulation becomes weaker and 
broader (w, and psi), and the precipitation, cloud fields, and 
circulation all consistently reside over the SST maximum (for about 1 year before going haywire).   
While the parameterized convection plays a large role in driving this asymmetry it appears to be not entirely a 
result of the convective parameterization, but also due to an interaction between the convective 
parameterization and the LW CRE.   The degree to which this asymmetry influences the comparison with other experiments is 
unclear.    Note: check the wind shear for P25large, P100small, and P100large (E25 has less shear than E1 or E2).

Complex patterns of precipitation over a fixed sinusoidal or Gaussian SST distribution have been noted many times in 
previous literature \cite{Bretherton2006, Wofsy2012} (Ming Zhao (personal communication, see Jeevanjee et al. 2018))
but the irregularities have 
tended to be symmetric about the SST maximum.  This is broadly consistent with our simulations when the convection is 
entirely explicit (E25, E2, and E1, discussed further in section 4).

%%%%%
% section 3.1 results
\begin{figure}
  \centering
      \includegraphics[width=0.96\columnwidth]{walkerfigs/rh_psi_cond_25km_6pan.eps}
      \caption{Comparison of the steady state relative humidity (shading of top panels) and mass 
      streamfunction (contours) for experiments
      with a grid-spacing of 25km on a domain of 200 $\times$ 4000 $\, \rm{km}^2$.  The lower panels 
      show the total (liquid + ice) condensate (g/kg).  
      An experiment with fully parameterized convection (P25) is on the left, the same 
      experiment but with the LWCRE off is in the center panel, and the experiment with LWCRE on, but the convective
      parameterization turned off is on the right (E25).  All three panels use the same contour interval for the mass 
      streamfunction (kg/s).}
  \label{fig:rh_psi_P25vsE25}
\end{figure}

@@

Previous studies have shown that cloud radiative effects act to strengthen and contract an overturning circulation 
(e.g. Popp and Silvers, 2017, HH, and others (Muller and Held?), Albern et al., 2019).  This is 
due to an increased low-level flux
of moist static energy into the convective regions.  While our mock-Walker circulation is distinct from radiative
convective equilibrium and the resulting convective self-aggregation, there are obvious similarities between our 
region of persistent deep convection and a state of aggregation.   One of the simplest measures of convective 
aggregation and the large-scale circulation is the fraction of the domain in which the air is subsiding, the 
subsidence fraction (SF).  As convection becomes more organized, or aggregated, SF will increase.  We expect 
for an overturning circulation that a contraction of the convective region would result in a larger subsidence 
fraction.  This is precisely what we see in Table 2.  For each of our experiments, the cases with LWCRE on
have a larger SF (with the exception of E1, for which SF is constant).  Our prescribed SST warm patch ensures that our simulations will 
be `aggregated' to some degree.  However, the subsidence fraction is a useful metric even in this case which 
includes an additional constraint than RCE.  Given identical SSTs, the different SFs give a measure of variability
that is driven entirely by the atmosphere.

@@

Despite the same boundary conditions and model base, the experiments documented here have a
large range of domain mean precipitation ($\rm{\overline{P}}$, Table 2) that varies by as much as 
0.6 $\rm{mm/d}$ (3.5-4.1 in parameterized experiments; 3.1-3.7 in explicit experiments).
This highlights how dominant the interaction between clouds and radiation can be in determining the 
characteristics of a system.
Because of the tight constraints that connect the domain mean precipitation, atmospheric condensational heating, and 
the total radiative cooling, the time evolution of the precipitation serves as a useful measure of whether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that this equilibrium is reached after about 30 
days for the E2, and E1 simulations, and after about 50 days for the P25 and P100 simulations.   After the initial adjustments
the simulations all oscillate about mean precipitation values which tend to increase with the grid-spacing 
(Table 2).   The fact that E1 and E2 reach equilibrium sooner than E25, P25, or P100 and that the period of oscillation 
about the domain mean precipitation is smaller helps to justify the 6 month simulation times for E1 and E2.  
%The period of oscillation is much larger for the GCM like simulations than it is for the CRM like oscillations.
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
\cite{Silvers2016, Patrizio2019}
%(Silvers et al., 2016; Patrizio and Randall, 2019).   
%The experiments with a higher resolution have a lower value of domain
%mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  
Differences in $\rm{\overline{P}}$ can be understood as a consequence of the differences in upper level cloud fraction 
and the surface energy budget and will be discussed further in a later section.

@@

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are 
two of the most fundamental choices that must be made.   One the goals of this study was to explore the results of 
changing grid-spacing over a wide range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies 
\cite{Bretherton2005, Bretherton2006, Muller2012, Jeevanjee2013, Silvers2016, Patrizio2019}
(e.g. Bretherton et al., 2005; Bretherton et al., 2006; Muller and Held, 2012; Jeevanjee and Romps, 2013; 
Silvers et al., 2016; Patrizio and Randall, 2019) have 
documented sensitivities of the equilibrated state to domain size.   
The analysis of the previous, and of the next sections focus on results from 
experiments using a domain with a long domain length of $4000 \, \rm{km}$.  However we do find interesting 
sensitivities to the domain size that are noted here.

The evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, 
and the effect of the longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a domain with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ 
(middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with previous work showing that CREs 
strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation changes more as a function of domain size than it does as a function of resolution. 
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitation.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.  As previously notes, these
GCM-like simulations equilibrate after approximately 50 days.  However, there is also a dramatic change in 
the distribution of precipitation on the $4000\, \rm{km}$ domain simulations after almost 2 years.   The domain mean 
precipitation does not significantly change in these cases, only the spatial structure.

@@

Smaller domains were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the small domains, relative to a control domain size that was 4x
larger.   In contrast, here larger domains have larger precipitation rates (Table 2).  For the LWCRE off cases, in agreement 
with \citeA{Bretherton2006} we find a more 
focused ascent region in the smaller domain.   But when the LWCRE is on, the extent of the ascent region changes little between P25 
and P25L but for P100 and P100L the ascent region is much more focuses on the large domain.    
How do our low-level clouds over the colder SSTs compare with the findings of B06 (check the condensate for small vs. large 
domains)?

% Fraction of Precip due to large-scale precip for large/small experiments with a 25km resolution
\begin{figure}
  \centering
      \includegraphics[width=0.8\columnwidth]{walkerfigs/Conv_vs_LS_hist.eps}
  \caption{Precipitation that is due to the large-scale(blue) cloud scheme and to the convective 
  parameterization (black).  Experiments with the LW CRE off are shown with thin lines.  
  Panels on the left show large domains with a width of 16000 km and
  panels on the right show domains with a width of 4000 km.}
  \label{fig:conv_vs_ls}
\end{figure}

The domain mean total precipitation is constrained by the radiative cooling of the atmosphere.  However, in 
models with the convection parameterized the 
total precipitation is composed of precipitation from the convection scheme and the large-scale cloud scheme. 
While the relative contribution of each component is not well constrained, it is thought to influence the 
atmospheric state (Held et al., what part of the state?). 
The distribution of convective and large-scale precipitation indicates how the condensational heating
in a GCM is being distributed among the parameterizations, and what is triggering the precipitation.  
Precipitation from each of these two components is
shown in Figure \ref{fig:conv_vs_ls} as a function of both resolution and domain size.  
In the regions of large-scale ascent, most of the 
precipitation derives from the large-scale cloud scheme.  We also see that the LW CRE (thick lines) dramatically increases
the large-scale precipitation.  The LW CRE has a much smaller effect on the magnitude of the convective precipitation 
but does act to spatially concentrate it.  With the exception 
of the P100L LW CRE off experiment, the convective precipitation produces relatively little of the total precipitation for 
both resolutions and on both domains.  The dramatic dependence on domain size of the precipitation field that is 
seen in Figure \ref{fig:domdep_lwoff} corresponds to a decrease in the large-scale precipitation of about 65\% in P25L case
and an almost complete elimination of the large-scale precipitation in the P100L case.

@@

% this paragraph should perhaps be moved to the introduction...
It is clear that the overturning circulation in the tropical Pacific encompasses dynamical motions at a host of scales and that the 
different cloud types interact with the circulation on scales ranging from meters to thousands of km's.  
The tropical Pacific also serves as an ideal location to study interactions between clouds and the 
circulation because it includes strong overturning circulations, abundant shallow cumulus, congestus, and cumulonimbus 
clouds (Johnson et al. 1999), as well as stratocumulus cloud decks along the eastern extremities of the basin.
Can we hope to simulate all of this in a single idealized modeling context?    
Our initial motivation for using the GFDL AM4 model on a doubly-periodic domain was to simulate 
a tropical Pacific-like region 
%that is rich in phenomena 
in the context of both a GCM and a CRM in the hopes that resolving more of the turbulent motions and circulations
would help us to better understand the physics and the mechanisms which are at work in the cloud-circulation
interactions of the tropical Pacific and improve our ability to model this region in a GCM.    
The dynamical core of the current generation GFDL models is ideally suited to this task because it can easily 
be used over a wide range of resolutions and can solve either the hydrostatic or nonhydrostatic 
governing equations.

\begin{figure}
  \centering
      \includegraphics[width=0.96\columnwidth]{walkerfigs/rh_psi_cond_E25E2E1.eps}
      \caption{The equilibrated state of the Walker cell as a function of resolution.  
      Experiments shown are E25 (left), E2(center), and E1(right). Top panels show the 
      steady state relative humidity (shading) and mass streamfunction (black contours)
      while bottom panels show the total condensation (liquid + ice).  
      All panels use the same contour interval for the 
      mass streamfunction (kg/s).}
  \label{fig:rh_psi_P25E2E1}
\end{figure}

@@

In this section we focus on simulations with grid-spacing of 1km, 2km, and 25km.   
A glance at Figure \ref{fig:hov_4mods_6mn} shows notable differences in the precipitation structure that result from the 
overturning circulation at different resolutions.  Shown are 180 days of precipitation from the P25, E25, E2, and 
E1 simulations.  As the 
resolution increases towards $1 \, \rm{km}$ the distribution of precipitation becomes more regular and consistently 
centered over the SST maxima in the middle of the domain.  Both the P25 and E25 simulations show more 
variability at later times compared to these first 180 days.  The simulations with explicit convection at resolutions 
typical of cloud-resolving models (E2, E1) show little aversion to the precipitation maximum occurring over 
the maximum in SST and relative to the P100, P25 and E25 simulations 
they are able to maintain a smoother distribution of  precipitation over a broader range of SST values.

@@

\bibliography{Silvers_WalkerCell}
